graphsim: An R package for simulating gene expression data from graph structures of biological pathways

### 2020-05-01 paper/paper.Rmd
correc figure formatting for graph structures

plot_directed(graph, state ="activating", layout = layout.kamada.kawai,
              cex.node = 2, cex.arrow = 4, arrow_clip = 0.2,
              sub = "(a) Activating pathway structure", cex.sub = 1.5)

@@

plot_directed(graph, state=state, layout = layout.kamada.kawai,
              cex.node = 2, cex.arrow = 4, arrow_clip = 0.2,
              sub = "(b) Inhibiting pathway structure", cex.sub = 1.5)
```

@@

```{r simulation_activating, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a graph structure}. An example of a graph structure (a) that has been used to derive a relationship matrix (b), $\\Sigma$  matrix (c) and correlation structure (d) from the relative distances between the nodes. Non-negative values are coloured white to red from $0$ to $1$. This $\\Sigma$ matrix has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution. Here genes with closer relationships in the pathway structure show a higher correlation between simulated values.'}
# activating graph
state <- rep(1, length(E(graph)))
plot_directed(graph, state=state, layout = layout.kamada.kawai,
              cex.node=3, cex.arrow=5, arrow_clip = 0.2,
              sub = "(a) Activating pathway structure", cex.sub = 1.5)

#adjacency matrix
adj_mat <- make_adjmatrix_graph(graph)
#relationship matrix
dist_mat <- make_distance_graph(graph, absolute = FALSE)
#sigma matrix directly from graph
sigma_mat <- make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE)
#show shortest paths of graph
shortest_paths <- shortest.paths(graph)
#generate expression data directly from graph
expr <- generate_expression(100, graph, cor = 0.8, mean = 0, comm = FALSE,
                            dist = TRUE, absolute = FALSE, state = state)
#plot adjacency matrix
heatmap.2(make_adjmatrix_graph(graph), scale = "none", trace = "none",
          col = colorpanel(3, "grey75", "white", "blue"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "Adjacency matrix", cex.sub = 1.5)
#plot relationship matrix
heatmap.2(make_distance_graph(graph, absolute = FALSE),
          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "(b) relationship matrix", cex.sub = 1.5)
#plot sigma matrix
heatmap.2(make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE),
scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = expression(paste("(c) ", Sigma, " matrix")), cex.sub = 1.5)
expr <- generate_expression(100, graph, cor = 0.8, mean = 0,
comm = FALSE, dist =TRUE, absolute = FALSE, state = state)
#plot simulated expression data
heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "(e) Simulated expression data (log scale)", cex.sub = 1.5)
#plot simulated correlations
heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
```

@@

#generate parameters for inhibitions
state <- c(1, 1, -1, 1, 1, 1, 1, -1)
plot_directed(graph, state=state, layout = layout.kamada.kawai,
              cex.node=3, cex.arrow=5, arrow_clip = 0.2,
              sub = "(a) Inhitibing pathway structure", cex.sub = 1.5)

#adjacency matrix
adj_mat <- make_adjmatrix_graph(graph)
#relationship matrix
dist_mat <- make_distance_graph(graph, absolute = FALSE)
#sigma matrix directly from graph
sigma_mat <- make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE)
#show shortest paths of graph
shortest_paths <- shortest.paths(graph)
#generate expression data directly from graph
expr <- generate_expression(100, graph, state = state, cor = 0.8, mean = 0, comm = FALSE,
                            dist = TRUE, absolute = FALSE)
#plot adjacency matrix
heatmap.2(make_adjmatrix_graph(graph), scale = "none", trace = "none",
          col = colorpanel(3, "grey75", "white", "blue"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "Adjacency matrix", cex.sub = 1.5)
# #plot relationship matrix
heatmap.2(make_distance_graph(graph, absolute = FALSE),
          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "(b) relationship matrix", cex.sub = 1.5)
# #plot sigma matrix
heatmap.2(make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE),
scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = expression(paste("(c) ", Sigma, " matrix")), cex.sub = 1.5)
expr <- generate_expression(100, graph,  state = state, cor = 0.8, mean = 0,
comm = FALSE, dist =TRUE, absolute = FALSE)
#plot simulated expression data
heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "(e) Simulated expressinon data (log scale)", cex.sub = 1.5)
#plot simulated correlations
heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
```

@@

plot_directed(graph, state=state, layout = layout.kamada.kawai,
              cex.node=3, cex.arrow=5, arrow_clip = 0.2,
              sub = expression(paste("(a) TFG-", Beta, " activates SMADs")), cex.sub = 1.5)

#adjacency matrix
adj_mat <- make_adjmatrix_graph(graph)
#relationship matrix
dist_mat <- make_distance_graph(graph, absolute = FALSE)
#sigma matrix directly from graph
sigma_mat <- make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE)
#show shortest paths of graph
shortest_paths <- shortest.paths(graph)
#generate expression data directly from graph
expr <- generate_expression(100, graph, state = state, cor = 0.8, mean = 0, comm = FALSE,
                            dist = TRUE, absolute = FALSE)
#plot adjacency matrix
heatmap.2(make_adjmatrix_graph(graph), scale = "none", trace = "none",
          col = colorpanel(3, "grey75", "white", "blue"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "Adjacency matrix", cex.sub = 1.5)
# #plot relationship matrix
heatmap.2(make_distance_graph(graph, absolute = FALSE),
          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "(b) relationship matrix", cex.sub = 1.5)
# #plot sigma matrix
heatmap.2(make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE),
scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = expression(paste("(c) ", Sigma, " matrix")), cex.sub = 1.5)
expr <- generate_expression(100, graph,  state = state, cor = 0.8, mean = 0,
comm = FALSE, dist =TRUE, absolute = FALSE)
#plot simulated expression data
heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "(e) Simulated expressinon data (log scale)", cex.sub = 1.5)
#plot simulated correlations
heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
```

>>>>

plot_directed(graph, state ="activating", layout = layout.kamada.kawai,
              cex.node = 2, cex.arrow = 4, arrow_clip = 0.2)
mtext(text = "(a) Activating pathway structure", side=1, line=3, at=0.05, adj=0.5, cex=1.75)
box()

@@

plot_directed(graph, state=state, layout = layout.kamada.kawai,
              cex.node = 2, cex.arrow = 4, arrow_clip = 0.2)
mtext(text = "(b) Inhibiting pathway structure", side=1, line=2, at=0.075, adj=0.5, cex=1.75)
box()
```

@@

```{r simulation_activating, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a graph structure}. An example of a graph structure (a) that has been used to derive a relationship matrix (b), $\\Sigma$  matrix (c) and correlation structure (d) from the relative distances between the nodes. Non-negative values are coloured white to red from $0$ to $1$. This $\\Sigma$ matrix has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution. Here genes with closer relationships in the pathway structure show a higher correlation between simulated values.'}
# activating graph
state <- rep(1, length(E(graph)))
plot_directed(graph, state=state, layout = layout.kamada.kawai,
              cex.node=2, cex.arrow=4, arrow_clip = 0.2,
              sub = "(a) Activating pathway structure", cex.sub = 1.75)
box()

#adjacency matrix
adj_mat <- make_adjmatrix_graph(graph)

#relationship matrix
dist_mat <- make_distance_graph(graph, absolute = FALSE)

#sigma matrix directly from graph
sigma_mat <- make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE)

#show shortest paths of graph
shortest_paths <- shortest.paths(graph)

#generate expression data directly from graph
expr <- generate_expression(100, graph, cor = 0.8, mean = 0, comm = FALSE,
                            dist = TRUE, absolute = FALSE, state = state)

#plot adjacency matrix
heatmap.2(make_adjmatrix_graph(graph), scale = "none", trace = "none",
          col = colorpanel(3, "grey75", "white", "blue"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "Adjacency matrix", cex.sub = 1.75)

#plot relationship matrix
heatmap.2(make_distance_graph(graph, absolute = FALSE),
          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "(b) relationship matrix", cex.sub = 1.75)

#plot sigma matrix
heatmap.2(make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE),
scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = expression(paste("(c) ", Sigma, " matrix")), cex.sub = 1.75)

#plot simulated expression data
expr <- generate_expression(100, graph, cor = 0.8, mean = 0,
comm = FALSE, dist =TRUE, absolute = FALSE, state = state)
heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "(e) Simulated expression data (log scale)", cex.sub = 1.75)

#plot simulated correlations
heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
```

@@

#generate parameters for inhibitions
state <- c(1, 1, -1, 1, 1, 1, 1, -1)
plot_directed(graph, state=state, layout = layout.kamada.kawai,
              cex.node=2, cex.arrow=4, arrow_clip = 0.2,
              sub = "(a) Inhitibing pathway structure", cex.sub = 1.75)
box()

#adjacency matrix
adj_mat <- make_adjmatrix_graph(graph)

#relationship matrix
dist_mat <- make_distance_graph(graph, absolute = FALSE)

#sigma matrix directly from graph
sigma_mat <- make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE)

#show shortest paths of graph
shortest_paths <- shortest.paths(graph)

#generate expression data directly from graph
expr <- generate_expression(100, graph, state = state, cor = 0.8, mean = 0, comm = FALSE,
                            dist = TRUE, absolute = FALSE)

#plot adjacency matrix
heatmap.2(make_adjmatrix_graph(graph), scale = "none", trace = "none",
          col = colorpanel(3, "grey75", "white", "blue"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "Adjacency matrix", cex.sub = 1.75)

# #plot relationship matrix
heatmap.2(make_distance_graph(graph, absolute = FALSE),
          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "(b) relationship matrix", cex.sub = 1.75)

# #plot sigma matrix
heatmap.2(make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE),
scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = expression(paste("(c) ", Sigma, " matrix")), cex.sub = 1.75)

#plot simulated expression data
expr <- generate_expression(100, graph,  state = state, cor = 0.8, mean = 0,
comm = FALSE, dist =TRUE, absolute = FALSE)
heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "(e) Simulated expression data (log scale)", cex.sub = 1.75)

#plot simulated correlations
heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
```

@@

plot_directed(graph, state = state, layout = layout.kamada.kawai,
              border.node=scales::alpha("black", 0.75), fill.node="lightblue",
              col.arrow = c(scales::alpha("navyblue", 0.25), scales::alpha("red", 0.25))[state], 
              cex.node = 1.5, cex.label = 0.8, cex.arrow = 2, 
              sub = expression(paste("(a) TFG-", Beta, " activates SMADs")), cex.sub = 1.75)
box()

#adjacency matrix
adj_mat <- make_adjmatrix_graph(graph)

#relationship matrix
dist_mat <- make_distance_graph(graph, absolute = FALSE)

#sigma matrix directly from graph
sigma_mat <- make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE)

#show shortest paths of graph
shortest_paths <- shortest.paths(graph)

#generate expression data directly from graph
expr <- generate_expression(100, graph, state = state, cor = 0.8, mean = 0, comm = FALSE,
                            dist = TRUE, absolute = FALSE)

#plot adjacency matrix
heatmap.2(make_adjmatrix_graph(graph), scale = "none", trace = "none",
          col = colorpanel(3, "grey75", "white", "blue"),
          colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "Adjacency matrix", cex.sub = 1.75)

# #plot relationship matrix
heatmap.2(make_distance_graph(graph, absolute = FALSE),
          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "(b) relationship matrix", cex.sub = 1.75)

# #plot sigma matrix
heatmap.2(make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE),
scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = expression(paste("(c) ", Sigma, " matrix")), cex.sub = 1.75)

#plot simulated expression data
expr <- generate_expression(100, graph,  state = state, cor = 0.8, mean = 0,
comm = FALSE, dist =TRUE, absolute = FALSE)
heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "(e) Simulated expression data (log scale)", cex.sub = 1.75)

#plot simulated correlations
heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
```
